### 2024-09-06 base-rates.qmd
20240906 - base rates

In sum, the [marginal probability](#sec-baseRate), including the [prior probability](#sec-baseRate) or [base rate](#sec-baseRate), should be weighed heavily in predictions unless there are sufficient data to indicate otherwise, i.e., to update the posterior probability based on new evidence.
Bayes' theorem provides a powerful tool to anchor predictions to the [base rate](#sec-baseRate) unless sufficient evidence changes the posterior probability (by updating the evidence and [prior probability](#sec-baseRate)).

>>>>

In sum, the [marginal probability](#sec-baseRate), including the [prior probability](#sec-baseRate) or [base rate](#sec-baseRate), should be weighed heavily in predictions unless there are sufficient data to indicate otherwise, i.e., to update the posterior probability based on new evidence.
Bayes' theorem specifies how prior beliefs (i.e., [base rate](#sec-baseRate) informations) should be integrated with the [predictive accuracy](#sec-predictiveValidity) of the evidence to make predictions.
It thus provides a powerful tool to [anchor](#sec-heuristicsAnchoringAdjustment) predictions to the [base rate](#sec-baseRate) unless sufficient evidence changes the posterior probability (by updating the evidence and [prior probability](#sec-baseRate)).
In general, you should [anchor](#sec-heuristicsAnchoringAdjustment) your predictions to the [base rate](#sec-baseRate) and [adjust](#sec-heuristicsAnchoringAdjustment) from there.
As noted by @Kahneman2011, if you have doubts about the quality of the evidence for a particular prediction question, keep your predictions close to the [base rate](#sec-baseRate), and modify them only modifying them based on the new information.

### 2024-09-06 cognitive-bias.qmd
20240906 - base rates

The *representativess heuristic* refers to the tendency for a person's judgments or predictions about individuals to be made based on how similar the individual is to (i.e., how closely the individual resembles) the person's existing mental prototypes.
For instance, when coming out of college, Tight End Kyle Pitts [drew comparisons](https://247sports.com/article/kyle-pitts-lebron-james-2021-nfl-draft-florida-gators-football-163882176) to the "LeBron James" of Tight Ends (archived at <https://perma.cc/JQB5-XPVL>).
The idea that his athletic profile leads him to be similar to the prototype of LeBron James may have led him to be too highly drafted by fantasy managers in his first seasons.

@@

The *base rate fallacy* (aka base rate neglect) is the tendency to ignore information about the general probability of an event in favor of specific information about the event.
The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
In general, people tend to overestimate the likelihood of low base-rate events.

>>>>

The *representativess heuristic* refers to the tendency for a person's judgments or predictions about individuals to be made based on how similar the individual is to (i.e., how closely the individual resembles) the person's existing mental prototypes.
In this way, the representativeness heuristic involves stereotyping, which is not a statistically optimal approach to predictions [@Kahneman2011].
Moreover, the representativness heuristic does not adequately consider the quality of the evidence [@Kahneman2011].
It is thus important to question the [predictive accuracy](#sec-predictiveValidity) of the evidence for the particular prediction question.

Consider the following example of the representativeness heuristic.
When coming out of college, Tight End Kyle Pitts [drew comparisons](https://247sports.com/article/kyle-pitts-lebron-james-2021-nfl-draft-florida-gators-football-163882176) to the "LeBron James" of Tight Ends (archived at <https://perma.cc/JQB5-XPVL>).
The idea that his athletic profile leads him to be similar to the prototype of LeBron James may have led him to be too highly drafted by fantasy managers in his first seasons.

@@

The *base rate fallacy* (aka base rate neglect) is the tendency to ignore information about the general probability of an event in favor of specific information about the event.
The *base rate* is a marginal probability, which is the general probability of an event irrespective of other things.
Focusing on the specific information about the event (and thus underweighting the base rate) involves the [representativeness heuristic](#sec-heuristicsRepresentativeness) [@Kahneman2011]—making judgments based on how similar the event is to a mental prototype regardless of how frequent the event the event actually is.
In general, people tend to overestimate the likelihood of low base-rate events [@Kahneman2011].
